\section{Summary}

\section{Future Work}

\begin{itemize}
\item Continue the implementation of the binary first non-trivial elimination
search to replace the working but inefficient linear search.
\item Implement a more efficient method for date emulation in the web
application.
\item Restructure design of desktop based and web based user interfaces to 
take into account constructive user interface design suggestions gained from 
the user evaluation process. 
 \end{itemize}

\section{Lessons Learned}

\subsection{Team communication}

At points during the course of the project, there has been significant 
communication break downs. Although the team decided to use numerous ways to 
communicate (Facebook, Google Docs, GitHub tickets, team meetings, and advisor 
meetings) smaller groupings of the team experienced significant breaks in 
communication. This was most apparent during important spells in the 
development cycle, for instance the Christmas break. Unfortunately some 
messages between team members were left unheard due to different team members 
focusing communication efforts on different mediums. To combat this, and using 
knowledge gained through experience, the team gradually reduced the amount of 
communication mediums it used as the project progressed. For future projects, 
the team would recommend keeping the amount of communication mediums used by 
to a minimum.      

\subsection{Team structure}

At the start of the project the team decided on a Laissez-Faire management 
structure, which it was thought would let individual team members gravitate 
towards areas of work that suited their strengths. However, at the end of the 
first semester it was apparent that this was not a suitable management 
structure due to some team members having a far greater technical capability, 
personal knowledge and effective work rate than other members of the team. It 
was decided that the team should restructured. One team meeting after the 
Christmas break provided a suitable time for the team to vote on a Leader and 
Second in Command. It was decided by the team that Gordon Reid would be 
selected for the role of team leader and Ryan Wells for the role of Second in 
Command. For any future projects the team would recommend using a top down 
management structure from the start of the project. This would allow more 
technically experienced and capable time-managing team members' skills to be 
drawn upon. This would also highlight members of the team who may 
unfortunately take advantage of not having an explicit leader by not 
contributing to the work flow to an acceptable level.           

\subsection{Distribution of work load}

As mentioned above, a sizeable concern for the duration of the project was the 
distribution of work load between team members. During the first few weeks of 
the development cycle it became clear that one or two members of the team had 
measurably greater technical capability and effective work rates than other 
team members. These team team members (One of whom who was chosen by the team 
as Team Leader, and the other as Second in Command), were invaluable during 
the course of the applications production life cycle, and again for the 
forward progression of the application as a finished product. Although there 
must be a small skill gap between all team members in any team development 
environment, usually this can be overcome by time expenditure, personal 
development and a good work ethic by team members who are less experienced. 
Unfortunately during the development of the application, at numerous times, 
multiple team members had to be pulled away from their current role to assist 
one team member who it was felt universally was not pulling their weight, or 
contributing to a satisfactory level. The team felt they dealt with this in a 
suitably professional and timely manner, usually during numerous discussions 
with the project supervisor.           

\subsection{Measuring Scope}

Prior to starting the implementation of the project, multiple functional
requirements were stated using the MoSCoW approach (Must have, Should have,
Could have, Won't have (but Would like)). The team decided to `play it safe'
and have few `must haves' to keep the scope of the project within a definitely
manageable size. This led to the must have functionality being implemented
and demonstrable shortly after the Christmas break. For future projects it may
be beneficial to expand the scope of the project to challenge and spur on
the entire team to take on more work.

\subsection{Timing of evaluation}

The user evaluation itself was felt by the team to be a huge success, effectively communicating numerous positive aspects of the design of the system, and also shed light on aspects of the system which needed revising. Although the team gathered numerous pieces of constructive feedback from the user evaluation, the team unfortunately did not have time to consider implementing any design changes suggested by the user evaluation due to the late point during the project life cycle that the user evaluation took place. If the team was to do the project again, it would have started the user evaluation process a few weeks earlier, to leave room at the end of the project life cycle to potentially implement any constructive design suggestions gained from the user evaluation process.