\subsection{Design}

\subsubsection{Introduction}

This section discusses the design of the web-based version of the application.
The design of the application was constructed with only the most important
functional and non-functional requirements in mind. The web-based version was
classed as the least important part of the project, with preliminary effort
directed solely towards the desktop application.

\subsubsection{System architecture}

The web application is a standard multi-tier architecture with the presentation,
logic, and data separated from each other.

The presentation tier is the client/browser who has Hyper Text Mark-up Language
(HTML) and Cascading Style Sheets (CSS) for the static presentation of content.
In addition there is JavaScript supported by JQuery JQueryUI for the dynamic
user interface elements.

The logic tier runs on a web server called Lighttpd (pronounced lighty) that is
supported by PHP: Hypertext Preprocessor (PHP). The logic tier has two data
sources that make up the data tier, a MySQL database containing the latest data
and a Java jar for looking back at older data.

The N-Tier Architecture diagram is available from appendix
\ref{fig:nTierArchitecture}.

\subsubsection{User interface}

The user interface of the web application was intended on being as close to the
desktop interface as viable within the constraints of a web browser and within
the realms of what is typical layout of a web page.

A wireframe for the web application is shown in figure
\ref{fig:webApplicationWireframe}.

The web application has a single page containing the six available divisions.
Each division is a table and only one is available for viewing. The reasoning
behind this is to keep as much information `about the fold' (above the lower
page boundary on a browser's window).

There are links at the top of each page that will allow the user to traverse
the entire date range for the season allowing them to view the scoreboard and
elimination status at any point in time.

\subsection{Implementation}

\subsubsection{Introduction}

This section discusses the implementation of the web-based version of the
application. The implementation discussion will be split up into the three
main sections as shown in figure \ref{fig:nTierArchitecture}.

\subsubsection{Presentation}

HTML/CSS/jQuery

\subsubsection{Logic}

PHP

\subsubsection{Data}

SQL + JAR
